\documentclass[a4paper,11pt]{article}

%%% ---------------- %%%
%%% --- Packages --- %%%
%%% ---------------- %%%
    \usepackage[utf8]{inputenc}
    \usepackage{authblk}
    \usepackage{graphicx}
    \usepackage[space]{grffile}
    \usepackage{latexsym}
    \usepackage{textcomp}
    \usepackage{longtable}
    \usepackage{tabulary}
    \usepackage{booktabs,array,multirow}
    \usepackage{amsfonts,amsmath,amssymb}
    \usepackage[table]{xcolor}
    \definecolor{lightgray}{gray}{0.9}
    \definecolor{lightergray}{gray}{0.7}
    \usepackage{colortbl}
    
    % macro for arbitray vertical line in tables 
    % see https://tex.stackexchange.com/questions/31441/vertical-line-only-in-one-part-of-a-table
    \def\arrvline{\hfil\kern\arraycolsep\vline\kern-\arraycolsep\hfilneg}
    % -- me 
    \usepackage{lineno}         % for line numbers 
    \usepackage{natbib}         % for citations
    \usepackage[a4paper, left = 2.5cm, right = 2.5cm, top = 2.5cm, bottom = 2.5cm]{geometry}            % page layout
    \usepackage{setspace}       % line spacing
    \usepackage{mathptmx}       % Times new Roman 1
    \usepackage[T1]{fontenc}    % Times new Roman 2
    \usepackage{subcaption}     % subfigure environment
    \usepackage{hyperref}
    \usepackage[labelfont=bf]{caption} % Labels dick gedruckt
    \usepackage[font=small]{caption}  % caption text body small
    \usepackage{microtype}          % to avoid over- or underfull hboxes 
    

    
%%% ---------------- %%%
%%% --- Settings --- %%%
%%% ---------------- %%%
\linenumbers        % line numbers
\doublespacing      % double spaced
\setlength\parindent{0pt}

 % --- Footnotes without marks --- % 
 % see https://tex.stackexchange.com/questions/30720/footnote-without-a-marker
 \newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}


% colors 


\title{Should ecologists prefer model- over algorithm-based multivariate methods?}

\author[1]{Jonathan F. Jupke}
\author[1]{Ralf B. Schäfer}
\affil[1]{iES Landau, Institute for Environmental Sciences, University Koblenz-Landau, Fortstraße 7, 76829 Landau, Germany}%

\date{}

\begin{document}

%%% ----------------- %%%
%%% ---Title Page --- %%%
%%% ----------------- %%%

    \maketitle

    %%% ---------------- %%%
    %%% --- Abstract --- %%%
    %%% ---------------- %%%

        \begin{abstract}%  Less than 250 words 
        %motivation 
            Ecological data sets often record the abundance of several species, together with a set of explanatory variables. 
            %
            Multivariate statistical methods are optimal to analyze such data and are thus frequently used in ecology for exploration, visualization, and inference. 
            %
            Most of the common approaches are algorithmic, they follow mathematical routines like correspondence - or distance-based analyses, which stands in stark contrast to univariate statistics, where data models, assuming specific distributions, are the norm.
            %
            However, through advances in statistical theory and computational power, models for multivariate data have gained traction. 
            %
        %problem statement
            Systematic and simulation-based performance evaluations of these methods are important as guides for practitioners but still lacking.
        %approach
            Here, we compare two model-based methods, multivariate Generalized Linear Models (MvGLM) and Constrained Quadratic Ordination (CQO), with two algorithm-based methods, distance-based redundancy analysis (dbRDA) and Canonical Correspondence Analysis (CCA).
            % 		
            We studied the performance of the methods to discriminate between causal variables and noise variables for 180 simulated data sets covering different sample sizes and data distributions.
        %results
            MvGLM and dbRDA differentiated accurately between causal and noise variables.
            %
            The former had the lowest false positive rate (0.008) while the latter had the lowest false negative rate (0.017).
            %CCA
            CQO and CCA had the highest false negative rate (0.291) and false positive rate (0.256), respectively. 
        %conclusions
           Our study shows that both model- and algorithm-based methods have their place in the ecologists statistical toolbox. 
           % 
        MvGLM and dbRDA are reliable for analysing species-environment relations, whereas both CQO and CCA exhibited considerable flaws especially with linear environmental gradients. 
           %
           
        \end{abstract}

    %%% ----------------- %%%
    %%% --- Key-Words --- %%%
    %%% ----------------- %%%

        {\bf Keywords:} 
        %1.
        statistical models,
        %2.
        numerical simulations,
        %3 
        ordination, 
        %4
        multivariate analysis, 
        %5
        variable selection
    \newpage
    %% -- why this study and your paper deserves to be honored as a Highlighted Student Paper -- %% 
        \textit{Commonly used multivariate methods in ecology might be too restrictive.} 
        %
        \textit{Alternatives exist but generalizable comparisons are lacking.}
        %
        \textit{We offer such testing and an actionable guide to multivarite.}
        
    %% -- Author Contributions -- %%    
        \blfootnote{Author Contributions: JFJ and RBS conceived the experiment. JFJ conducted the simulation and the analyses. JFJ and RBS wrote the manuscript.}
        \blfootnote{Corresponding Author: Jonathan F. Jupke; jonathanfrederik@aol.com}
    
    
    \newpage
    %%% ------------- %%%
    %%% --- Parts --- %%%
    %%% ------------- %%%
\newpage
\input{01_Introduction}
\input{02_Methods}
\input{03_Results}
\input{04_Discussion}

%%% -------------------- %%%
%%% --- Bibliography --- %%%
%%% --------------------- %%%
\bibliographystyle{apalike}
\bibliography{references}

\end{document}
